Hierarchical Region-Network Sparsity for\\ High-Dimensional Inference in Brain Imaging

### 2016-02-05 paper/icml2016/example_paper.tex
interm

\twocolumn[
\icmltitle{Region-Network Hierarchical Sparsity Priors for\\
High-Dimensional Inference in Brain Imaging}
% \icmltitle{Region-Network Hierarchical Sparsity Priors for\\
% High-Dimensional Inference in Brain Imaging}

@@

\begin{abstract} 
Structured sparsity penalization has recently improved
statistical models applied to high-dimensional data in various domains.
As an extension to imaging neuroscience,
the present work incorporates
priors on network hierarchies of brain regions into
logistic-regression estimators to distinguish neural activity.
This enables recombination of
the neurobiological architectures underlying
functional segregation into brain regions and
functional integration by brain networks
that are routinely studied separately in neuroscientific research.
%
Region-network hierarchical priors for classifying
18 psychological tasks
from a reference dataset are
shown to outperform naive $\ell_1$-norm, elastic-net, and
trace-norm penalization, as well as neurobiologically
informed $\ell_1/\ell_2$-block-norm group sparsity.
Weighing the relative importance of
region and network structure within the hierarchical tree penalty
recovered complementary aspects of the neural activity patterns.
%
It is thus demonstrated
that priors of local and global neurobiological knowledge
can enhance
generalization performance, sample complexity, and domain interpretability
by hierarchical tree sparsity.

@@

% \paragraph{Problem formulation.}
\subsection{Problem formulation}
We formulate our estimation problem in the framework of regularized risk
estimation applied to linear models: 
The goal is to estimate a good predictor of cognitive task
given a brain image. Let the set \(\mathcal X\subset\R^p\) represent brain
images of \(p > 0\) voxels.
%
We minimize the risk \(\mathcal L(\hat y, y)\) with
\(\hat y = f(X\hat w + \hat b)\), where \(f\) is a link function 
(e.g. sigmoid for logistic regression, identity for linear regression),
and \(\mathcal L\) usually represents an appropriate negative loglikelihood.
We incorporate a useful prior through regularization:
\[\argmin_{w, b} \mathcal L(f(Xw + b), y) + \lambda\Omega(w),\]
where \(\lambda > 0\) and \(\Omega\) is the regularizer.

Brain \textit{regions} are defined as disjoint groups of voxels. Let \(\mathcal G\)
be a partition of \(\{1, \dots, p\}\), i.e.
\[  \bigcup_{i} g_i = \{1, \dots, p\} \textrm{ and } g_i\cap g_j=\emptyset
\quad\forall i\not=j\]

Brain \textit{networks} consist of a group of one or several brain regions.
The set of brain networks \(\mathcal{H}\) also forms a partition of 
\(\{1, \dots, p\}\) and in addition it is consistent with \(\mathcal G\) in
the sense that
\[\textrm{for all }g\in\mathcal G, h\in\mathcal H,\quad
\textrm{ either } g\subset h\textrm{ or }g\cap h = \emptyset.\]
This allows a clear association of each region \(g\in\mathcal G\) to a 
network \(h\in\mathcal H\) and thus establishes a tree structure (up to 
adding a root node containing all voxels).

>>>>

\twocolumn[
\icmltitle{Hierarchical Region-Network Sparsity for\\
High-Dimensional Inference in Brain Imaging}
% \icmltitle{Region-Network Hierarchical Sparsity Priors for\\
% High-Dimensional Inference in Brain Imaging}

@@

\begin{abstract} 
Structured sparsity penalization has recently improved
statistical models applied to high-dimensional data in various domains.
%
As an extension to imaging neuroscience,
the present work incorporates
priors on network hierarchies of brain regions into
logistic-regression to distinguish neural activity.
This recombines neurobiological architectures underlying
functional segregation into brain regions and
functional integration by brain networks
that are studied separately in neuroscientific research.
%
Hierarchical region-network priors are shown to better classify
18 psychological tasks from a reference dataset
than other sparse estimators.
Weighing the relative importance of
region and network structure within the hierarchical tree penalty
recovered complementary aspects of the neural activity patterns.
%
Local and global priors of neurobiological knowledge
are thus demonstrated to enhance
generalization performance, sample complexity, and domain interpretability.

@@

% \paragraph{Problem formulation.}
\subsection{Problem formulation}
We formulate our estimation problem in the framework of regularized risk
estimation applied to linear models: 
The goal is to estimate a good predictor of cognitive task
given a brain image. Let the set \(\mathcal X\subset\R^p\) represent brain
images of \(p > 0\) voxels.

Then we would like to minimize the risk \(\mathcal L(\hat y, y)\), where
\(\hat y = X\hat w + \hat b\), while regularizing to incorporate a useful
prior:
\[\argmin_{w, b} \mathcal L(Xw + b, y) + \lambda\Omega(w),\]
where \(\lambda > 0\) and \(\Omega\) is the regularizer.

Brain regions are defined as disjoint groups of voxels. Let \(\mathcal G\)
be a partition of \(\{1, \dots, p\}\), i.e.
\[  \bigcup_{i} g_i = \{1, \dots, p\} \textrm{ and } g_i\cap g_j=\emptyset
\quad\forall i\not=j\]

Brain networks consist of regions and are thus super-regions or groups of 
regions. The set of brain networks \(\mathcal{H}\) is also a partition of 
\(\{1, \dots, p\}\) and in addition it is consistent with \(\mathcal G\) in
the sense that
\[\textrm{for all }g\in\mathcal G, h\in\mathcal H,\quad
\textrm{ either } g\subset h\textrm{ or }g\cap h = \emptyset.\]
This allows a clear association of each region \(g\in\mathcal G\) to a 
network \(h\in\mathcal H\) and thus establishes a tree structure (up to 
adding a root node containing all voxels).
